Co-citations in context

### 2019-06-12 cocit_r.tex
just a bunch of copyediting

\cite{uzzi_atypical_2013} used a novel approach for co-citation analysis to characterize a subset of highly cited articles with respect to both novel and conventional combinations of prior research. The frequency with which references were co-cited  in 17.9 million articles and their cited references from the Web of Science (WoS) was calculated and expressed as journal pair frequencies (observed co-citation frequencies). Expected co-citation values were generated from Monte Carlo simulations under a random graph model. Observed frequencies were then normalized (shifted and scaled) to averaged expected values from ten simulations and termed as \emph{z-scores}. Consequently, every article was associated with multiple z-scores corresponding to co-cited journal pairs in its references. For each article, positional statistics of z-scores were calculated to set thresholds for a binary (high/low) classification of conventionality (C) using the median z-score of an article and novelty (N) using the tenth percentile of z-scores within an article. Thus, HCLN would denote high conventionality and low novelty with all four combinations being possible. The authors observed that HCHN articles were twice as likely to be highly cited, suggesting that novel combinations of ideas flavoring a body of conventional thought were a feature of impact and that `science follows a nearly universal pattern'.

Key to Uzzi {\em et al.}, is their random graph model and its underlying assumptions. The citation switching algorithm used to generate expected values from random substitutions is designed to preserve the number of publications, the number of references in each publication, the year of publication of both publications and references, and the disciplinary composition of the references cited in these publications. However, substitutions are equiprobable with respect to disciplinary origin and citation count. The model permits substitution of a reference in quantum physics with equal probability by a reference in quantum physics, quantum chemistry, classical literature, entomology, or anthropology. Such substitutions poorly model the disciplinary nature of scientific endeavor and citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which appears inconsistent with the power law or lognormal citation distributions described in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  Accordingly, model mis-specification is likely to arise on account of the simulated values not modeling the empirical data very well.

@@

Despite different methods used to generate expected values, both of these key preceding studies (\emph{vide supra}) measured co-citation frequencies across the scientific literature (the WoS superset) and normalized them without disciplinary constraints before subsequently analyzing disciplinary subsets.
In other words, the concerns raised above about mis-specification resulting from treating
all references as equiprobable, independent of discipline, apply to \cite{boyack_vs_uzzi_2014} as well.
 We hypothesized that modifying the normalization to constrain references to be drawn only from the disciplinary citation network,   rather than all of WoS, would reduce model mis-specification.  Consequently, we used keyword searches of the scientific literature to construct exemplar citation networks themed around academic disciplines: \emph{applied physics, immunology, and metabolism}. Within these disciplinary frameworks, we calculated observed and expected co-citation frequencies using a refined random graph model and an efficient Monte Carlo simulation algorithm.

Our analyses, using three different techniques, provides substantial evidence that constraining the reference substitutions to the disciplinary subnetwork reduces model mis-specification compared to treating all substitutions within WoS as equiprobable. 
Furthermore, re-analyses of these three semantically-themed
 citation networks  under the improved model reveals strikingly different trends than those 
 observed by Uzzi {\em et al.} 
 For example, while Uzzi {\em et al.}~claimed that highly cited articles are more likely to be both HC and HN than expected, and that this trend held across all subdisciplines, we find that these trends vary with the subdiscipline so that no universal trends can be established. Specifically, 
 HC remains highly correlated with highly cited articles in the immunology and metabolism datasets but not with applied physics, and HN is highly correlated with highly cited articles in applied physics but not with immunology and metabolism.    
 Thus, disciplinary networks are different from each other, and trends that hold for the full WoS network do not hold for even large subnetworks (such as metabolism). 
 Furthermore, we also found that the categories HC, HN, etc.,  demonstrating the highest percentage of highly cited articles are not robust with respect to varying thresholds for high citation counts or for highly novel citation patterns.
 Overall, our study, although limited to three subnetworks, suggests that  co-citation analysis, while useful in many regards, cannot be used to provide universal features in scholarship, due to important differences between disciplines and between publications.

\section{Materials and Methods}
\emph{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 using data in ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073), which were then stratified by year of publication. For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(Table \ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data (Table \ref{tab:summary_data}), we analyzed the two boundary years (1985 and 2005) and the mid-point (1995). We also used the number of times each of these articles was cited in the first 8 years since publication as a measure of its impact.
\vspace{-4mm}
% latex table generated in R 3.5.3 by xtable 1.8-4 package
% Fri May 17 14:28:31 2019
\begin{table}[ht]
\caption{Summary of base WoS Analytical Dataset.
The number of unique publications of type Article, unique references (UR), total references (TR), and the ratio of total references to unique references increases monotonically with each year indicating that both the number of documents and citation activity increase over time. Only publications of type Article with at least two references and references with complete publication data were included.} 
\label{tab:summary_data}
\centering
\scalebox{0.7}{
\begin{tabular}{|r  cccc |}
  \hline
year & Unique Publications & Unique References (UR) & Total References (TR) & TR/UR \\ 
  \hline
1985 & 391,860 & 2,266,584 & 5,588,861 & 2.47 \\ 
  \hline
1986 & 402,309 & 2,316,451 & 5,708,796 & 2.46 \\ 
1987 & 412,936 & 2,427,347 & 5,998,513 & 2.47 \\ 
1988 & 426,001 & 2,545,647 & 6,354,917 & 2.50 \\ 
1989 & 443,144 & 2,673,092 & 6,749,319 & 2.52 \\ 
1990 & 458,768 & 2,827,517 & 7,209,413 & 2.55 \\ 
1991 & 477,712 & 2,977,784 & 7,729,776 & 2.60 \\ 
1992 & 492,181 & 3,134,109 & 8,188,940 & 2.61 \\ 
1993 & 504,488 & 3,278,102 & 8,676,583 & 2.65 \\ 
1994 & 523,660 & 3,458,072 & 9,255,748 & 2.68 \\ 
  \hline
1995 & 537,160 & 3,680,616 & 9,875,421 & 2.68 \\ 
  \hline
1996 & 663,110 & 4,144,581 & 11,641,286 & 2.81 \\ 
1997 & 677,077 & 4,340,733 & 12,135,104 & 2.80 \\ 
1998 & 693,531 & 4,573,584 & 12,728,629 & 2.78 \\ 
1999 & 709,827 & 4,784,024 & 13,280,828 & 2.78 \\ 
2000 & 721,926 & 5,008,842 & 13,810,746 & 2.76 \\ 
2001 & 727,816 & 5,203,078 & 14,261,189 & 2.74 \\ 
2002 & 747,287 & 5,464,045 & 15,001,390 & 2.75 \\ 
2003 & 786,284 & 5,773,756 & 16,024,652 & 2.78 \\ 
 2004 & 826,834 & 6,095,594 & 17,167,347 & 2.82 \\ 
   \hline
 2005 & 886,648 & 6,615,824 & 19,036,324 & 2.88 \\ 
 \hline
\end{tabular}}
\vspace{-4mm}%Put here to reduce too much white space after your table 
\end{table}

We constructed three disciplinary datasets based on keyword searches. (i) immunology (ii) metabolism (iii) applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(Table \ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched traditional subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science; life sciences \& biomedicine, physical sciences, technology, social sciences, arts \& humanities using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.
\vspace{-2 mm}
\begin{table}[ht]
\caption{Disciplinary Datasets. PubMed and WoS were searched for articles using search terms, `immunology', `metabolism', and `applied physics'. Counts of publications are shown for each of the three years analyzed and expressed in parentheses as a percentage of the total number of publications in our analytical WoS dataset (Table \ref{tab:summary_data}) for that year.}
\label{tab:disc-datasets}
\centering
\begin{tabular}{|r r r r|}
  \hline
Year & Immunology & Metabolism & Applied Physics \\ 
  \hline
1985 & 21,606 (5.5\%) & 78,998 (20.2\%) & 10,298 (2.7\%) \\ 
1995 & 29,320 (5.5\%)  & 121,247 (22.6\%) & 21,012 (3.9\%)  \\ 
2005 & 37,296 (4.2\%) & 200,052 (22.6\%) & 35,600 (4.0\%)  \\ 
 \hline
\end{tabular}
\vspace{-6mm}
\end{table}

\emph{Monte Carlo simulations, normalization of observed frequencies, annotations, and `hit' papers}. Analysis was performed on publications from 1985, 1995, and 2005. Building upon prior work~\citep{uzzi_atypical_2013}, all ${n \choose 2}$ reference pairs were generated for each publication, where $n$ is the number of cited references in the publication. These reference pairs were then mapped to the journals they were published in using ISSN numbers as identifiers. Where multiple ISSN numbers exist for a journal, the most frequently used one in the WoS was assigned to the journal. In addition, publications containing fewer than two references were discarded. Journal pair frequencies were summed across the dataset to create observed frequencies $(F_{obs})$.

@@

We invoke model mis-specification due to considering disciplinarily irrelevant references as a concern when disciplinary subsets are analyzed using a random graph model that generates expected co-citation frequencies from the WoS superset. 
A symptom of model misspecification is that random sampling of citations associated with an article do not match with the distribution of citations across the disciplines observed in articles from that discipline. To address this issue, we first assembled three exemplar datasets using keyword searches for the terms `applied physics', `immunology', and `metabolism' (Materials and Methods). Using the WoS extended classification of 153 subject areas, we then analyzed the disciplinary composition of the references in these three datasets into the five research areas in the Web of Science (life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities) for the years 1985, 1995, and 2005. We observed that 84.2\% of the references in the applied physics dataset were classified as belonging to the WoS physical sciences research area, which is comparable to 79.5\% for the references in the physical sciences area. Similarly, 93.7\% and 92.1\% of the references in the immunology and metabolism datasets belonged to the life sciences and biomedicine research area, again comparable to 89.9\% for the references in the life sciences and biomedicine area. \emph{Add in Sitaram's new data}. These data indicate strongly that references are likely to belong to the same research area as the publications that cite them. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, although that percentage varied among disciplines in the eight specialties they investigated: from approximately 39\% to 89\% for 2006.

@@

We predicted that model misspecification, as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies, constraining  reference substitutions to the disciplinary subnetwork would be less than the
model misspecification when substitutions are drawn equiprobably from all of 
the WoS superset. 
The results in Table \ref{tab:kld} indicate that for the set of journals common to both a disciplinary network and the WoS superset,  confirm this prediction: simulations under our model (which constrains substitutions to the
disciplinary subnetwork) consistently have a lower K-L divergence compared to simulations that draw from the WoS superset. 
Furthermore, the K-L divergence for  the unconstrained model (where the references are equiprobable and allowed to be taken from all of WoS) is generally twice as large as the K-L divergence for the models where the references are constrained to  the disciplinary subnetworks (a ratio of 1.96 to 2.77,  and  greater than 2.0 for eight out of nine cases), demonstrating that simulations that constrain substitutions to the given disciplinary network better model the observed data.

@@

Figure \ref{fig:Fig2}, Panels (a) and (b), compares hit rates for the four categories among the Immunology, Applied Physics, and WoS datasets for 1995: the hit rate is defined as the number of hit articles in each category divided by the number of articles in the category. We evaluated the statistical significance of the categorical hit rates using multiple methods, some of which we describe here.  Our first test was based on the null hypotheses that hits were distributed randomly among the four categories with uniform probability in proportion to the number of articles in each category. Using a Chi-Square Goodness of Fit test, rejecting the null hypothesis in favor of the alternate hypothesis supports a non-uniform dispersion of hits: that is, some of the four categories are individually associated with higher than expected, or lower than expected hit rates. The null hypothesis was rejected at a $p<0.001$ in all cases in  Figure \ref{fig:Fig2}, with the exception of the immunology and applied datasets where hit articles are designated as the top 1\% of articles: valid tests were not possible in those instances due to too few expected hits. The null hypothesis was rejected with $p<0.001$ for all valid tests for all parameter settings, all datasets, and all years: hypotheses tests were valid in 73 of 96 instances. We conclude that it is likely that the distribution of hits among categories is not uniform but, rather, hit rates vary among the categories in all datasets. (Should we insert a table with all statistical results?)

We computed hit rates for the WoS dataset, which mirrored Uzzi et al.'s results whereby the largest hit rates were for the HNHC category, despite our methodological improvement of sampling citations in proportion to their frequency. We found contrary results in the 1995 immunology dataset where articles in the LNHC category often had the greatest hit rates, as reflected in Table \ref{Tab:HitRate} and Figure \ref{fig:Fig2}.  Across all year's data, all datasets, and all parameter settings, the highest hit rates in the immunology datasets were sometimes in the LNHC category and sometimes in the HNHC category. The metabolism hit rates reflected this same pattern. The greatest hit rates for the applied physics data were often in the HNLC category as reflected in Table \ref{Tab:HitRate} and Figure \ref{fig:Fig2}, and otherwise in the HNHC category. We conclude that Uzzi et al.'s finding of high hit rates in the HNHC category does not hold generally for disciplinary-based datasets and that novel citation patterns are not always indicative of impactful research, as was the case with immunology. Furthermore, the categories displaying the greatest hit rate vary with parameter settings and with the year. The lack of stable results across parameter settings suggests that parameters must be selected judiciously.

We also tested the explanatory power of each framework dimension by classifying articles as Low Novelty (LN) or High Novelty (HN) and, separately, as having Low Conventionality (LC) or High Conventionality (HC).  We tested the null hypothesis that hits are distributed between LN and HN (LC and HC) in proportion to the total number of articles assigned to those categories.  That null hypothesis was rejected for the WoS data along both dimensions. Consistent with previous analysis, hit articles were overrepresented in the HC category in every instance of WoS data at a $p<0.001$ and hit articles were overrepresented in the HN category at a $p<0.001$ in all but two cases.  The p-values, in those cases, were 0.002 and 0.007.  Hits in the immunology and metabolism data were overrepresented in the HC category with the same statistical significance as for WoS. The relationship of novelty with hits in the immunology and metabolism data differed dramatically from the WoS, however, with statistically significant findings of hit articles being sometimes overrepresented in the LN category, and sometimes being underrepresented.  Of the 12 tests for applied physics, the statistical significance supporting a positive relationship between hit articles and HN were all $p<0.10$, and 10 of 12 were $p<0.05$.  These tests also indicated strong support for the relationship between LN and hit articles in applied physics in a limited number of tests, with $p<0.10$ in 5 of 12 instances and $p < 0.05$ in 3 of 12 instances. These results suggest that (1) both conventionality and novelty are strongly related to hits in the WoS, (2) the conventionality dimension is strongly related with hits in immunology and metabolism and novelty is not, (3) novelty is more strongly related with hits in applied physics than is conventionality. More generally, we find that the dimensions most strongly related with hit articles vary across disciplines and between disciplinary and broad data sets.

@@

\emph{needs to be expanded}
We conclude that high hit rates in the HNHC category do not hold generally for all datasets and that conventionality flavored with novelty is not always indicative of impactful research. 
%The z-scores that change sign across datasets is either contradictory or an acceptable variation due to the difference in reference sets. We contend that it is the former because the inappropriate substitution of citations with those from disciplines that are implausible and because the observed citation frequencies are ignored. 
That fewer z-scores are negative in the WoS dataset relative to the immunology dataset may be directly due to the uniform sampling of references, whereby many resulting journal pairs are never observed in the literature and so the expected frequencies of those observed pairs have lower expected values, thus biasing their z-scores upward. In addition, treating citation as a set versus a multiset means that a random model will sample popular citations downward, thus further increasing their z-scores.\emph{need confirmation from Jim here}.

>>>>

\cite{uzzi_atypical_2013} used a novel approach for co-citation analysis to characterize a subset of highly cited articles with respect to both novel and conventional combinations of prior research. The frequency with which references were co-cited  in 17.9 million articles and their cited references from the Web of Science (WoS) was calculated and expressed as journal pair frequencies (observed co-citation frequencies). Expected co-citation values were generated from Monte Carlo simulations under a random graph model. Observed frequencies were then normalized (shifted and scaled) to averaged expected values from ten simulations and termed as \emph{z-scores}. Consequently, every article was associated with multiple z-scores corresponding to co-cited journal pairs in its references. For each article, positional statistics of z-scores were calculated to set thresholds for a binary (high/low) classification of conventionality (C) using the median z-score of an article and novelty (N) using the tenth percentile of z-scores within an article. Thus, HCLN would denote high conventionality (HC) and low novelty (LN), with all four combinations being possible. The authors observed that HCHN articles were twice as likely to be highly cited, suggesting that novel combinations of ideas flavoring a body of conventional thought were a feature of impact and that `science follows a nearly universal pattern'.

Key to Uzzi {\em et al.}, is their random graph model and its underlying assumptions. The citation switching algorithm used to generate expected values from random substitutions is designed to preserve the number of publications, the number of references in each publication, the year of publication of both publications and references, and the disciplinary composition of the references cited in these publications. However, a critically important feature of the model is that all references published in the same year have equal probability of being cited by any given publication.
Thus, in particular,  subject matter and  citation count do not impact the probability that the reference will be cited, so that  a reference in quantum physics can be substituted, with equal probability, by a reference in quantum physics, quantum chemistry, classical literature, entomology, or anthropology. Such substitutions poorly model the disciplinary nature of scientific endeavor and citation behavior~\citep{wallace_lariviere_gingras_2012,moed_measuring_2010,klavans_research_2017,garfield_1979}. In addition, under this random model, a reference cited over 100 times in a given year is selected with the same probability as a reference cited only once, which appears inconsistent with the power law or lognormal citation distributions described in the literature~\citep{stringer_statistical_2010,perline_strong_2005}.  Accordingly, model misspecification is likely to arise on account of the simulated values not modeling the empirical data very well.

@@

Despite different methods used to generate expected values, both of these key preceding studies (\emph{vide supra}) measured co-citation frequencies across the scientific literature (the WoS superset) and normalized them without disciplinary constraints before subsequently analyzing disciplinary subsets.
In other words, the concerns raised above about misspecification resulting from treating
all references as equiprobable, independent of discipline, apply to \cite{boyack_vs_uzzi_2014} as well.
 We hypothesized that modifying the normalization to constrain references to be drawn only from the disciplinary citation subnetwork,   rather than all of WoS, would reduce model misspecification.  Consequently, we used keyword searches of the scientific literature to construct exemplar citation networks themed around academic disciplines: \emph{applied physics, immunology, and metabolism}. Within these disciplinary frameworks, we calculated observed and expected co-citation frequencies using a refined random graph model and an efficient Monte Carlo simulation algorithm.

Our analyses, using multiple techniques, provides substantial evidence that constraining the reference substitutions to the disciplinary subnetwork reduces model misspecification compared to treating all substitutions within WoS as equiprobable. 
Furthermore, re-analyses of these three semantically-themed
 citation networks  under the improved model reveals strikingly different trends than those 
 observed by Uzzi {\em et al.} 
 For example, while Uzzi {\em et al.}~claimed that highly cited articles are more likely to be both HC and HN than expected, and that this trend held across all subdisciplines, we find that these trends vary with the subdiscipline so that no universal trends can be established. Specifically, 
 HC remains highly correlated with highly cited articles in the immunology and metabolism datasets but not with applied physics, and HN is highly correlated with highly cited articles in applied physics but not with immunology and metabolism.    
 Thus, disciplinary networks are different from each other, and trends that hold for the full WoS network do not hold for even large subnetworks (such as metabolism). 
 Furthermore, we also found that the categories HC, HN, etc.,  demonstrating the highest percentage of highly cited articles are not robust with respect to varying thresholds for high citation counts or for highly novel citation patterns.
 Overall, our study, although limited to three disciplinary subnetworks, suggests that  co-citation analysis, while useful in many regards, cannot be used to provide universal features in scholarship, due to important differences between disciplines and between publications.

\section{Materials and Methods}

\subsection{Bibliographic data} We have previously developed ERNIE, an open source knowledge platform into which we parse the Web of Science (WoS) Core Collection~\citep{Keserci371955}. WoS data stored in ERNIE spans the period 1900-2019 and consists of over 72 million publications. For this study, we generated an analytical dataset from years 1985 to 2005 using data in ERNIE. The total number of publications in this dataset was just over 25 million publications (25,134,073), which were then stratified by year of publication. For each of these years, we further restricted analysis to publications of type Article. Since WoS data also contains incomplete references or references that point at other indexes, we also considered only those references for which there were complete records~(Table \ref{tab:summary_data}). For example, WoS data for year 2005 contained 1,753,174 publications, which after restricting to type Article and considering only those references described above resulted in 916,573 publications, 6,095,594 unique references (set of references), and 17,167,347 total references (multiset of references). Given consistent trends in the data (Table \ref{tab:summary_data}), we analyzed the two boundary years (1985 and 2005) and the mid-point (1995). We also used the number of times each of these articles was cited in the first 8 years since publication as a measure of its impact.
\vspace{-4mm}
% latex table generated in R 3.5.3 by xtable 1.8-4 package
% Fri May 17 14:28:31 2019
\begin{table}[ht]
\caption{Summary of base WoS Analytical Dataset.
The number of unique publications of type Article, unique references (UR), total references (TR), and the ratio of total references to unique references increases monotonically with each year indicating that both the number of documents and citation activity increase over time. Only publications of type Article with at least two references and references with complete publication data were included.} 
\label{tab:summary_data}
\centering
\scalebox{0.7}{
\begin{tabular}{|r  cccc |}
  \hline
year & Unique Publications & Unique References (UR) & Total References (TR) & TR/UR \\ 
  \hline
1985 & 391,860 & 2,266,584 & 5,588,861 & 2.47 \\ 
  \hline
1986 & 402,309 & 2,316,451 & 5,708,796 & 2.46 \\ 
1987 & 412,936 & 2,427,347 & 5,998,513 & 2.47 \\ 
1988 & 426,001 & 2,545,647 & 6,354,917 & 2.50 \\ 
1989 & 443,144 & 2,673,092 & 6,749,319 & 2.52 \\ 
1990 & 458,768 & 2,827,517 & 7,209,413 & 2.55 \\ 
1991 & 477,712 & 2,977,784 & 7,729,776 & 2.60 \\ 
1992 & 492,181 & 3,134,109 & 8,188,940 & 2.61 \\ 
1993 & 504,488 & 3,278,102 & 8,676,583 & 2.65 \\ 
1994 & 523,660 & 3,458,072 & 9,255,748 & 2.68 \\ 
  \hline
1995 & 537,160 & 3,680,616 & 9,875,421 & 2.68 \\ 
  \hline
1996 & 663,110 & 4,144,581 & 11,641,286 & 2.81 \\ 
1997 & 677,077 & 4,340,733 & 12,135,104 & 2.80 \\ 
1998 & 693,531 & 4,573,584 & 12,728,629 & 2.78 \\ 
1999 & 709,827 & 4,784,024 & 13,280,828 & 2.78 \\ 
2000 & 721,926 & 5,008,842 & 13,810,746 & 2.76 \\ 
2001 & 727,816 & 5,203,078 & 14,261,189 & 2.74 \\ 
2002 & 747,287 & 5,464,045 & 15,001,390 & 2.75 \\ 
2003 & 786,284 & 5,773,756 & 16,024,652 & 2.78 \\ 
 2004 & 826,834 & 6,095,594 & 17,167,347 & 2.82 \\ 
   \hline
 2005 & 886,648 & 6,615,824 & 19,036,324 & 2.88 \\ 
 \hline
\end{tabular}}
\vspace{-1mm}%Put here to reduce too much white space after your table 
\end{table}

We constructed three disciplinary datasets based on keyword searches: immunology,   metabolism, and   applied physics. For the first two, rooted in biomedical research, we searched Pubmed for the term `immunology' or `metabolism' in the years 1985, 1995, and 2005~(Table \ref{tab:disc-datasets}). Pubmed IDs (pmids) returned were matched to WoS IDs (wos\_ids) and used to retrieve relevant articles. For the applied physics dataset, we directly searched traditional subject labels in WoS for `applied physics'. We also examined publications in the five major research areas in the Web of Science: life sciences \& biomedicine, physical sciences, technology, social sciences, and arts \& humanities,  using the extended subcategory classification of 153 sub-groups to categorize disciplinary composition of cited references in the datasets we studied.
\vspace{-2 mm}
\begin{table}[ht]
\caption{Disciplinary Datasets. PubMed and WoS were searched for articles using search terms, `immunology', `metabolism', and `applied physics'. Counts of publications are shown for each of the three years analyzed and expressed in parentheses as a percentage of the total number of publications in our analytical WoS dataset (Table \ref{tab:summary_data}) for that year. Note that Applied Physics and Immunology each represent about 4\% of the publications in WoS, but Metabolism occupies nearly one-fourth of WoS.}
\label{tab:disc-datasets}
\centering
\begin{tabular}{|r r r r|}
  \hline
Year & Applied Physics & Immunology & Metabolism   \\ 
  \hline
1985 & 10,298 (2.7\%) & 21,606 (5.5\%) & 78,998 (20.2\%)  \\ 
1995 & 21,012 (3.9\%)  & 29,320 (5.5\%)  & 121,247 (22.6\%)   \\ 
2005 & 35,600 (4.0\%) & 37,296 (4.2\%) & 200,052 (22.6\%)    \\ 
 \hline
\end{tabular}
\vspace{-6mm}
\end{table}

\subsection{Monte Carlo simulations, normalization of observed frequencies, annotations, and `hit' papers}

We performed analyses on publications from 1985, 1995, and 2005. Building upon prior work~\citep{uzzi_atypical_2013}, all ${n \choose 2}$ reference pairs were generated for each publication, where $n$ is the number of cited references in the publication. These reference pairs were then mapped to the journals they were published in using ISSN numbers as identifiers. Where multiple ISSN numbers exist for a journal, the most frequently used one in the WoS was assigned to the journal. In addition, publications containing fewer than two references were discarded. Journal pair frequencies were summed across the dataset to create observed frequencies $(F_{obs})$.

@@

Recall our concern regarding model misspecification  in the Uzzi {\em et al.}~model arising from treating all references in a given year as equiprobable, thus allowing substitutions from WoS without consideration of subject matter (as reflected in the publication's discipline) or citation frequency.
Under this model, the probability of replacing a reference by another reference in a given discipline is identical to the proportion of the articles in WoS that are in that particular discipline.
Thus, if the model accurate reflects citation practice, the proportion of references within papers published in a given discipline $D$ would be the same as the proportion of articles in the discipline $D$, and--conversely--the degree to which the proportion deviates from the expected value reflects the degree of model misspecification.
%due to considering disciplinarily irrelevant references as a concern when disciplinary subsets are analyzed using a random graph model that generates expected co-citation frequencies from the WoS superset. 
%A symptom of model misspecification is that random sampling of citations associated with an article do not match with the distribution of citations across the disciplines observed in articles from that discipline. 
We therefore  assembled three exemplar datasets using keyword searches for the terms `applied physics', `immunology', and `metabolism' (Materials and Methods). 
Using the WoS extended classification of 153 subject areas, we then analyzed the disciplinary composition of the references in these three datasets into the five research areas in the Web of Science (life sciences and biomedicine, physical sciences, social sciences, technology, and arts and humanities) for the years 1985, 1995, and 2005. We observed that 84.2\% of the references in the applied physics dataset were classified as belonging to the WoS physical sciences research area, which is close to the value of 79.5\% for the references in the physical sciences area. Similarly, 93.7\% and 92.1\% of the references in the immunology and metabolism datasets belonged to the life sciences and biomedicine research area, again close to the value of 89.9\% for the references in the life sciences and biomedicine area. \emph{Add in Sitaram's new data}. 
These data indicate strongly that references are likely to belong to the same research area as the publications that cite them. Our observations are in agreement with \cite{wallace_lariviere_gingras_2012} who found that, often, a majority of an article's citations are from the specialty of the article, although that percentage varied among disciplines in the eight specialties they investigated: from approximately 39\% to 89\% for 2006. 
\textcolor{blue}{Would be nice to know the proportion of publications in each of the five major subject areas.}

@@

We predicted that model misspecification, as measured by the Kullback-Leibler (K-L) Divergence~\citep{kullback_information_1951} between observed and simulated frequencies, constraining  reference substitutions to the disciplinary subnetwork would be less than the
model misspecification when substitutions are drawn equiprobably from all of 
the WoS superset. 
The results in Table \ref{tab:kld} indicate that for the set of journals common to both a disciplinary network and the WoS superset,  confirm this prediction: simulations under our model (which constrains substitutions to the
disciplinary subnetwork) consistently have a lower K-L divergence compared to simulations that all replacements uniformly  from the WoS superset. 
Furthermore, the K-L divergence for  the unconstrained model (where the references are equiprobable and allowed to be taken from all of WoS) is generally twice as large as the K-L divergence for the models where the references are constrained to  the disciplinary subnetworks (a ratio of 1.96 to 2.77,  and  greater than 2.0 for eight out of nine cases), demonstrating that simulations that constrain substitutions to the given disciplinary network better model the observed data.

@@

Figure \ref{fig:Fig2}, Panels (a) and (b), compares hit rates for the four categories among the Immunology, Applied Physics, and WoS datasets for 1995, where the hit rate is defined as the number of hit articles in each category divided by the number of articles in the category. We evaluated the statistical significance of the categorical hit rates using multiple methods, some of which we describe here.  Our first test was based on the null hypotheses that hits were distributed randomly among the four categories with uniform probability in proportion to the number of articles in each category. Using a Chi-Square Goodness of Fit test, rejecting the null hypothesis in favor of the alternate hypothesis supports a non-uniform dispersion of hits: that is, some of the four categories are individually associated with higher than expected, or lower than expected hit rates. The null hypothesis was rejected at a $p<0.001$ in all cases in  Figure \ref{fig:Fig2}, with the exception of the immunology and applied datasets where hit articles are designated as the top 1\% of articles: valid tests were not possible in those instances due to too few expected hits. The null hypothesis was rejected with $p<0.001$ for all valid tests for all parameter settings, all datasets, and all years: hypotheses tests were valid in 73 of 96 instances. We conclude that it is likely that the distribution of hits among categories is not uniform and that instead hit rates vary among the categories in all datasets. 
%(Should we insert a table with all statistical results?)

We computed hit rates for the WoS dataset, which mirrored Uzzi et al.'s results whereby the largest hit rates were for the HNHC category, despite our methodological improvement of sampling citations in proportion to their frequency. We found contrary results in the 1995 immunology dataset where articles in the LNHC category often had the greatest hit rates (Figure \ref{fig:Fig2}).  
Across all years and parameter settings, the highest hit rates in the immunology datasets were sometimes in the LNHC category and sometimes in the HNHC category, and the hit rates for the metabolism datasets reflected this same pattern  (Figure \ref{fig:Fig2}). 
The greatest hit rates for the applied physics data were often in the HNLC category and otherwise in the HNHC category (Figure \ref{fig:Fig2}).
We conclude that Uzzi et al.'s finding of high hit rates in the HNHC category does not hold generally for disciplinary-based datasets and that novel citation patterns are not always indicative of impactful research, as was the case with immunology. 
Furthermore, the categories displaying the greatest hit rate vary with parameter settings and with the year. 
The lack of stable results across parameter settings suggests that parameters must be selected judiciously.

We also tested the explanatory power of each framework dimension by classifying articles as Low Novelty (LN) or High Novelty (HN) and, separately, as having Low Conventionality (LC) or High Conventionality (HC).  We tested the null hypothesis that hits are distributed between LN and HN (LC and HC) in proportion to the total number of articles assigned to those categories.  That null hypothesis was rejected for the WoS data along both dimensions. Consistent with previous analysis, hit articles were overrepresented in the HC category in every instance of WoS data at a $p<0.001$ and hit articles were overrepresented in the HN category at a $p<0.001$ in all but two cases.  
The p-values, in those cases, were $0.002$ and $0.007$.  
Hits in the immunology and metabolism data were overrepresented in the HC category with the same statistical significance as for WoS. The relationship of novelty with hits in the immunology and metabolism data differed dramatically from the WoS, however, with statistically significant findings of hit articles being sometimes overrepresented in the LN category, and sometimes being underrepresented.  Of the 12 tests for applied physics, the statistical significance supporting a positive relationship between hit articles and HN were all $p<0.10$, and 10 of 12 were $p<0.05$.  These tests also indicated strong support for the relationship between LN and hit articles in applied physics in a limited number of tests, with $p<0.10$ in 5 of 12 instances and $p < 0.05$ in 3 of 12 instances. These results suggest that (1) both conventionality and novelty are strongly related to hits in the WoS, (2) the conventionality dimension is strongly related with hits in immunology and metabolism and novelty is not, (3) novelty is more strongly related with hits in applied physics than is conventionality. More generally, we find that the dimensions most strongly related with hit articles vary across disciplines and between disciplinary and broad data sets.

@@

\emph{needs to be expanded}

We conclude that high hit rates in the HNHC category do not hold generally for all datasets and that conventionality flavored with novelty is not always indicative of impactful research. 
%The z-scores that change sign across datasets is either contradictory or an acceptable variation due to the difference in reference sets. We contend that it is the former because the inappropriate substitution of citations with those from disciplines that are implausible and because the observed citation frequencies are ignored. 
That fewer z-scores are negative in the WoS dataset relative to the immunology dataset may be directly due to the uniform sampling of references, whereby many resulting journal pairs are never observed in the literature and so the expected frequencies of those observed pairs have lower expected values, thus biasing their z-scores upward. In addition, treating citation as a set versus a multiset means that a random model will sample popular citations downward, thus further increasing their z-scores.\emph{need confirmation from Jim here}.
